# Lab 5: Mixed-Method Research

*Inequality and Democracy, Yet Again*

Return again to this regression:

```{r, results='hide'}
inequality <- read.csv("https://raw.githubusercontent.com/jnseawright/practice-of-multimethod/main/data/inequality.csv")

ineqlm2 <- lm(Gini ~ Polity + log(GDP), data=inequality)
summary(ineqlm2)
```

Last time, you conducted case studies of the Republic of Yemen and of
Zimbabwe. Using the data, determine which case-selection rule was used
to select these cases. Conduct one similar case study using surprising
causes case selection. The code below chooses the five top cases according to this rule.

```{r, results='hide'}
auxlm <- lm(Polity ~ log(GDP), data=inequality)
inequality$Country[order(abs(auxlm$residuals), decreasing=TRUE)[1:5]]
```


How do the insights generated by this case study
compare with those produced by yesterday's case studies?

*Case Selection in Practice*

For a research topic that interests you, find five published case
studies or multi-method studies with a significant case-study component.

How, if at all, is case selection described in these studies? What
advantages and disadvantages do you see in the case selection processes
used? Describe how you would select cases if you were to repeat these
studies, and explain why.
